---
title: "Chapter 5 Exercises"
author: "JC"
format: html
editor: visual
---

```{r}
library(rosdata)
library(tidyverse)
```

5.1)

```{r}
shooter <- function(rate = .6){
  i <- 0
  counter <- 0
  while(i < 2){
    counter <- counter + 1
    
    shot <- rbinom(1, size = 1, prob = rate)
    if(shot == 0){
      i <- i + 1
    }
    else{
      i <- 0
    }
  }
  return(counter)
}
```

b\)

```{r}
set.seed(10)
two_misses_straight <- replicate(n = 1000, shooter())
cat("mean =", mean(two_misses_straight), ", median =", median(two_misses_straight), ", sd =", sd(two_misses_straight), ", mad sd =", mad(two_misses_straight))
```

```{r}
hist(two_misses_straight)
```

c\)

```{r}
shooter_2 <- function(rate = .6){
  i <- 0
  counter <- 0
  shots_made <- 0
  while(i < 2){
    counter <- counter + 1
    
    shot <- rbinom(1, size = 1, prob = rate)
    if(shot == 0){
      i <- i + 1
    }
    else{
      shots_made <- shots_made + 1
      i <- 0
    }
  }
  return(c(shot_count = counter, shots_made = shots_made))
}
```

```{r}
set.seed(10)
two_misses_straight_2 <- replicate(n = 1000, shooter_2())
two_misses_straight_df <- as.data.frame(t(two_misses_straight_2))
head(two_misses_straight_df)
```

```{r}
two_misses_straight_df |> 
  ggplot(aes(x = shot_count, y = shots_made/shot_count)) +
  geom_point()
```

5.2)

```{r}
elevator_sim <- function(){
  male <- rbinom(10,1,.5)
  weight <- ifelse(male == 1,rnorm(1, 5.13, .17), rnorm(1, 4.96, .2))
  return(sum(exp(weight)))
}
```

```{r}
set.seed(10)
elevator_samples <- replicate(n = 10000,elevator_sim())
mean(elevator_samples > 1750)
```

The probability that the total weight exceeds this limit is 21.2%.

5.3)

a\)

```{r}
dbinom(x = 3, size = 10, prob = .4)
```

b\)

```{r}
three_shots <- function(){
  shots <- rbinom(n = 1, size = 10, prob = .4)
  return(sum(shots) == 3)
}
```

```{r}
set.seed(100)
three_shots_sims <- replicate(n = 10000,three_shots())
mean(three_shots_sims)
```

We see that the two are very close to each other.

5.4)

```{r}
twenty_sum <- function(){
  sims <- runif(n = 20,min = 0, max = 1)
  return(sum(sims))
}
```

```{r}
set.seed(10)
twenty_sum_sims <- replicate(n = 1000, twenty_sum())
hist(twenty_sum_sims, probability = TRUE, ylim = c(0,.4),
     breaks = 30)
curve(
  dnorm(x, mean = 20*.5, sd = sqrt(20 * (1/12))),
  add = TRUE,
  lwd = 2
)
```

The CLT estimate does a really good job.

5.5)

```{r}
height_diff <- function(){
  men <- rnorm(100, mean = 69.1, sd = 2.9)
  men_avg <- mean(men)
  women <- rnorm(100, mean = 63.7, sd = 2.7)
  women_avg <- mean(women)
  return(men_avg - women_avg)
}
```

```{r}
#set.seed(10)
height_diff_sim <- replicate(n = 1000, height_diff())
cat("mean =", mean(height_diff_sim)," sd = ",sd(height_diff_sim))
```

The exact values would be:

```{r}
cat("mean diff exact = ",69.1 - 63.7,"sd difference exact ",sqrt((2.9/10)^2 
                                                                 + (2.7/10)^2))
```

Which is in line with our simulation.

5.6)

```{r}
savings_sim <- function(){
  cost_savings <- rnorm(1, mean = 5, sd = 4)
  market_size <- rnorm(1, mean = 40000, sd = 10000)
  difference <- cost_savings*market_size
  return(difference)
}
```

```{r}
set.seed(10)
savings_sim_sampling <- replicate(n = 100000, savings_sim())
```

```{r}
mean(savings_sim_sampling)
sd(savings_sim_sampling)
```

```{r}
hist(savings_sim_sampling)
```

5.7)

```{r}
norm_sim <- function(){
  norm_sample <- rnorm(10, mean = 6, sd = 4)
  avg_value <- mean(norm_sample)
  lower_ci_50 <- avg_value + qnorm(c(.25))*sd(norm_sample)
  upper_ci_50 <- avg_value + qnorm(c(.75))*sd(norm_sample)
  lower_ci_95 <- avg_value + qnorm(c(.025))*sd(norm_sample)
  upper_ci_95 <- avg_value + qnorm(c(.975))*sd(norm_sample)
  return(
    c(
      avg_value    = avg_value,
      lower_ci_50  = lower_ci_50,
      upper_ci_50  = upper_ci_50,
      lower_ci_95  = lower_ci_95,
      upper_ci_95  = upper_ci_95
    )
  )
}
```

```{r}
set.seed(10)
norm_sim_sample <- replicate(n = 100, norm_sim())
norm_sim_sample_df <- as.data.frame(t(norm_sim_sample))
head(norm_sim_sample_df)
```

```{r}
norm_sim_sample_df <- norm_sim_sample_df |> 
  mutate(row_id = row_number())
norm_sim_sample_df |> 
  ggplot(aes(x = row_id, y = avg_value)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_ci_50, ymax = upper_ci_50)) +
  geom_errorbar(aes(ymin = lower_ci_95, ymax = upper_ci_95), alpha = .5) +
  geom_hline(aes(yintercept = 6))
```

5.8)

```{r}
experiment_sim <- function(){
  effect <- rnorm(127,mean = .1, sd = .17)
  avg_effect <- mean(effect)
  sd_effect <- sd(effect)
  lower_ci <- avg_effect + qnorm(.025)*sd_effect
  upper_ci <- avg_effect + qnorm(.975)*sd_effect
  interval_contains <- between(.1, lower_ci, upper_ci)
  return(
    c(
      avg_effect        = avg_effect,
      sd_effect         = sd_effect,
      lower_ci          = lower_ci,
      upper_ci          = upper_ci,
      interval_contains = interval_contains
    )
  )
}
```

```{r}
set.seed(10)
experiment_sim_sample <- replicate(n = 1000, experiment_sim())
experiment_sim_df <- as.data.frame(t(experiment_sim_sample))
head(experiment_sim_df)
```

b\)

```{r}
sum(experiment_sim_df$interval_contains)
```

c\)

```{r}
average_effect <- mean(experiment_sim_df$avg_effect)
sd_effect <- mean(experiment_sim_df$sd_effect)
average_effect;sd_effect
```

```{r}
set.seed(123)

# given values
true_theta <- 0.10
se <- 0.17
n_sims <- 1000

# (a) simulate 1000 point estimates
estimates <- rnorm(n_sims, mean = true_theta, sd = se)

# (b) construct 95% confidence intervals
lower <- estimates - 1.96 * se
upper <- estimates + 1.96 * se

# check coverage
covers_true <- lower <= true_theta & upper >= true_theta
num_cover <- sum(covers_true)
prop_cover <- mean(covers_true)

# (c) mean and sd of simulated estimates
mean_est <- mean(estimates)
sd_est <- sd(estimates)

# results
num_cover
prop_cover
mean_est
sd_est
```

5.9)

```{r}
significant_estimates <- estimates[lower > 0]
mean(significant_estimates);sd(significant_estimates)
```

We can see that if we are only looking at statistically significant evidence we have a massive effect size, which is far from the actual effect size of .1.

5.10)

```{r}
cost_effect_ratio_sim <- function(se_effectiveness = 1){
  cost_diff <- rnorm(1, 600, 400)
  effective_diff <- rnorm(1, 3, se_effectiveness)
  ratio <- cost_diff/effective_diff
  return(c(
    cost_diff = cost_diff,
    effective_diff = effective_diff,
    ratio = ratio
  ))
}
```

a\)

```{r}
set.seed(5)
cost_effect_sample_a <- replicate(1000, cost_effect_ratio_sim())
cost_effect_a_df <- as.data.frame(t(cost_effect_sample_a))
head(cost_effect_a_df)
```

```{r}
cost_effect_a_df |> 
  ggplot(aes(x = cost_diff, effective_diff)) +
  geom_point()
```

b\)

```{r}
avg_ratio <- mean(cost_effect_a_df$ratio);avg_ratio
ci_50 <- quantile(cost_effect_a_df$ratio,c(.25,.75));ci_50
ci_95 <- quantile(cost_effect_a_df$ratio,c(.025,.975));ci_95
```

c\)

```{r}
set.seed(10)
cost_effect_sample_c <- replicate(1000, 
                                  cost_effect_ratio_sim(se_effectiveness = 2))
cost_effect_c_df <- as.data.frame(t(cost_effect_sample_c))
head(cost_effect_c_df)
```

```{r}
cost_effect_c_df |> 
  ggplot(aes(x = cost_diff, effective_diff)) +
  geom_point()
```

```{r}
avg_ratio <- mean(cost_effect_c_df$ratio);avg_ratio
ci_50 <- quantile(cost_effect_c_df$ratio,c(.25,.75));ci_50
ci_95 <- quantile(cost_effect_c_df$ratio,c(.025,.975));ci_95
```

5.12)

```{r}
treatment_assigner <- function(n, p, code = "IRA"){
  if(n <= 0){
    stop("N must be greater than 0")
  }
  if(p<0|p>1){
    stop("Probability must be between 0 and 1 range")
  }
  if(code == "IRA"){
    return(rbinom(n,1,p))
  }
  if(code == "CRA"){
    if(n*p != floor(n*p)){
      warning("N*P is not an integer, 
              rounding to closest number to select number of treatments")
    }
    treatment_count <- round(n*p)
    control_count <- n - treatment_count
    samples <- sample.int(n, size = n)
    treatments <- as.integer(samples <= treatment_count)
    return(treatments)
  }
  if(code == "MP"){
    if(n%%2 != 0){
      stop("N must be divisible by two")
    }
    return(rep(c(1,0),n/2))
  }
  
}
```

```{r}
treatment_assigner(n = 11, p = .2, code = "CRA")
```

```{r}
treatment_assigner(n = 10, p = .5, code = "MP")
```
