---
title: "Chapter 6 Exercises"
author: "JC"
format: html
editor: visual
---

```{r}
library(rosdata)
library(tidyverse)
library(rstanarm)
```

6.7)

5.2)

```{r}
model_sim <- function(a = 1, b = 5, n = 10, sigma = 50){
  x_sim <- runif(n, min = 0, max = 100)
  errors <- rnorm(n, 0, sigma)
  y <- a + b*x_sim + errors
  fake <- data.frame(x_sim, y)
  fit <- stan_glm(y ~ x_sim, data = fake)
  print(fit, digits = 2)
  plot(x_sim, y)
  abline(fit)
}
```

```{r}
model_sim()
```

```{r}
model_sim(a = 100, b = 40, n = 200, sigma = 1000)
```

3\)

```{r}
model_sim(a = 100, b = 40, n = 20000, sigma = 1000)
```

Estimates approach true value and uncertainties decrease.

6.4)

```{r}
model_sim <- function(a = 1, b = 5, n = 10, sigma = 50){
  x_sim <- runif(n, min = 0, max = 100)
  errors <- rnorm(n, 0, sigma)
  y <- a + b*x_sim + errors
  fake <- data.frame(x_sim, y)
  fit <- stan_glm(y ~ x_sim, data = fake)
  return(fit)
}
```

```{r}
n_seq <- seq(10,100,10)
models <- lapply(n_seq, function(n) model_sim(a = 1, b = 5, n = n, sigma = 10))
```

```{r}
names(models) <- n_seq
summaries <- map2_dfr(models, names(models), \(mod,n_name){
  draws <- as.matrix(mod)
  tibble(
    n = n_name,
    term = colnames(draws),
    median = apply(draws, 2, median),
    mad_sd = apply(draws, 2, mad)
  )
})
```

```{r}
summaries <- summaries |> mutate(n = as.integer(n))
```

```{r}
ggplot(data = summaries,aes(x = n, y = median, col = term)) +
  geom_point() +
  geom_errorbar(aes(ymin = median - 2 * mad_sd, ymax = median + 2 * mad_sd))
```

The slope term is identified very early, as there is so much variation in x. Sigma and the intercept have intervals that shrink until about the n = 60 mark, as n increases our estimates get closer to the real values.

6.5)

```{r}
earnings <- read.csv("Data/ROS-Examples-master/Earnings/data/earnings.csv")
head(earnings)
```

```{r}
average_heights <- earnings |> 
  group_by(male) |> 
  summarize(avg_height = mean(height))
average_heights
```

```{r}
male_earnings <- -26 + .6*70.08889 + 10.6;male_earnings
female_earnings <- -26 + .6*64.48642;female_earnings

```

b\)

```{r}
avg_earnings <- .52*female_earnings + .48*male_earnings;avg_earnings
```

c\)

```{r}
sample_avg_earnings_by_gender <- earnings |> 
  group_by(male) |> 
  summarise(avg_earn = mean(earn))
sample_avg_earnings_by_gender
```

These are about 3k off our estimates.

```{r}
mean(earnings$earn)
```

About 2k off our estimate.

6.6)

```{r}
heights <- read.table("Data/ROS-Examples-master/PearsonLee/data/MotherDaughterHeights.txt", header = TRUE,sep = " ")
head(heights)
```

```{r}
avg_mother_height <- mean(heights$mother_height);avg_mother_height
```

```{r}
short_moms <- heights |> filter(mother_height < avg_mother_height);head(short_moms)
```

```{r}
fit <- stan_glm(daughter_height ~ mother_height, data = short_moms)
```

```{r}
print(fit)
```

We can see that the intercept of 30.4 and slope of .5 are very close to the intercept of 30 and slope of .54 when using the full dataset.

b\)

```{r}
avg_daughter_height <- mean(heights$daughter_height);avg_daughter_height
```

```{r}
short_daughters <- heights |> filter(daughter_height < avg_daughter_height)
head(short_daughters)
```

```{r}
fit <- stan_glm(daughter_height ~ mother_height, data = short_daughters)
```

```{r}
print(fit)
```

The intercept jumps a lot, and the slope is much smaller.

c\)

When we filter on
