---
title: "Exercises Chapter 2"
author: "JC"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(baseballr)
library(rosdata)
```

4.1)

```{r}
n <- 1000
n_treatment <- 500
n_control <- 500
prop_treatment <- .5
prop_control <- .4

sd_treatment <- sqrt(.5*(1 - .5)/n_treatment)
sd_control <- sqrt(.4*(1 - .4)/n_control)
prop_treatment;prop_control;sd_treatment;sd_control
```

```{r}
prop_diff <- prop_treatment - prop_control
prop_diff_se <- sqrt(sd_treatment^2 + sd_control^2)
prop_diff;prop_diff_se
```

4.2)

Let's assume that 55% of men vote for a candidate and 45% of women vote for a candidate. Then we have the standard deviations of $\sqrt{(.55)(1-.55)/n_{men}},\sqrt{(.45)(1 - .45)/n_{women})}$ respectively. That happens to be equal. Now to get a standard error of less than 5 percentage points we need:

$$
\sqrt{.55(1-.55)/{n_{men}}+.45(1-.45)/n_{women}}
$$

If we assume we can get an equal sample of men and women, we have:

$$
\sqrt{.55(1-.55)/n_{group}}=.05,.55(1-.55)/n_{group}=.05^2,.55,n_{group}=1/(.05^2(.55(1-.55)))=99
$$

```{r}
#prop_men <- .55
#prop_women <- .45
n_group <-  2*.55*(1 - .55)/(.05^2);
n_group
```

```{r}
se_men <- sqrt(.55*(1 -.55)/(n_group));se_men
se_women <- sqrt(.45*(1-.45)/(n_group));se_women
se_total <- sqrt(se_men^2 + se_women^2);se_total
```

4.3)

```{r}
p_1 <- .3
p_2 <- .4
n <- 20
r_1 <- rbinom(100000,20,p_1)
r_2 <- rbinom(100000,20,p_2)
mean(r_1 < r_2)
```

4.4)

Lets say we want a 95% chance of detecting, around n = 135.

```{r}
p_1 <- .3
p_2 <- .4
n <- 135
r_1 <- rbinom(100000,n,p_1)
r_2 <- rbinom(100000,n,p_2)
mean(r_1 < r_2)
```

4.6)

```{r}
girls <- read.table("Data/ROS-Examples-master/Girls/girls.dat",
           skip = 1)
```

a\)

Standard deviation in dataset:

```{r}
obs_sd <- sd(girls$V1); obs_sd
```

```{r}
avg_prop <- mean(girls$V1)
avg_prop
```

```{r}
exp_sd <- sqrt(avg_prop*(1 - avg_prop)/3900); exp_sd
```

The standard deviation is about what we would expect.

```{r}
diff <- obs_sd - exp_sd
diff
```

b\)

```{r}
test_stat <- obs_sd^2 *(24 - 1)/exp_sd^2
test_stat
pchisq(test_stat,df = 23)
```

A p-value of .096 indicates there is no significant difference between the observed and hypothesized value.

4.7)

```{r}
y <- 0;n <- 50
```

Basic CI:

```{r}
est <- y
se <- sqrt(est*(1 - est)/n)
se
```

```{r}
est + qnorm(c(.025,.975))*se
```

Jeffsonian CI:

```{r}
est_alt <- (y + 2)/(n + 4)
est_se <- sqrt(est_alt*(1 - est_alt)/(n + 4));
est_se
```

```{r}
est_alt + qnorm(c(.025,.975))*est_se
```

But we would really have the lower interval be zero.

4.8)

a\)

The point estimate would be $\log(1.42)$:

```{r}
log(1.42)
```

b\)

We know that the interval is about 1.96 SEs wide in each direction so:

```{r}
se <- (log(1.98) - log(1.42))/1.96;se
```

4.9)

WE have $\theta=p(1)+(1-p)\frac{1}{4}=1/4+3/4p$

Then since $\hat{\theta} = .6$:

```{r}
theta_hat = .6
p = 4/3*(theta_hat - 1/4);
p
```

And:

$$
SE(\hat{p})=\frac{4}{3}SE(\hat{\theta})
$$

```{r}
theta_se <- sqrt(theta_hat*(1 - theta_hat)/100);theta_se
p_se <- 4/3*theta_se
p + qnorm(c(.025,.975))*p_se
```

```{r}
n <- 100
prop <- .6
prop_guess <- .25
```

```{r}
prop_diff <- prop - prop_guess
prop_se <- sqrt(prop*(1 - prop)/n)
prop_guess_se <- sqrt(prop_guess*(1 - prop_guess)/n)
prop_diff_se <- sqrt(prop_se^2 + prop_guess_se^2);
prop_diff + c(-1.96,1.96)*prop_diff_se
```

4.10)\
a)

When assessing population estimates we know that:

$$
SE(\hat{p})=\sqrt{\hat{p}(1-\hat{p})/n}
$$

In which case the random sample of 1k will give us better population estimates.

However for comparison we have:

$$
SE(\hat{p}_{\text{diff}})=\sqrt{SE(p_{\text{latino}})^2+SE(p_{\text{non latino}})^2}
$$

For this it is likely that the sample of 300 latino with 700 non-latinos will perform better, as the error of the latino estimate will be much smaller.

b\)

```{r}
n_nl_a <- 850
n_l_a <- 150

n_nl_b <- 700
n_l_b <- 300
p <- .5
```

First we calculate the standard error of the population estimate. We have:

$$
\text{Pop}=.15\text{ Latino } +.85\text{Non-Latino}
$$

For the simple random sampling this is easy, as we are drawing directly from the population. So:

```{r}
se_pop_a <- sqrt(p*(1 - p)/(n_nl_a + n_l_a));se_pop_a
```

However for the population estimate for the stratified sample we need to first compute the error for the Latino and Non-Latino samples.

```{r}
se_l_b <- sqrt(p*(1 - p)/n_l_b)
se_nl_b <- sqrt(p*(1 - p)/n_nl_b)
se_l_b;se_nl_b
```

Then we have:

$$
\text{SE(Pop)}=.15 \text{ SE(Latino) }+.85 \text{ SE(Non-Latino)}
$$

So:

```{r}
se_pop_b <- .15*se_l_b + .85 * se_nl_b; se_pop_b
```

We can see that the standard error of the population estimate is larger for the stratified sample.

Now for the comparison we compute:

$$
\sqrt{\text{SE(Latino)}^2+\text{SE(Non-Latino)}^2}
$$

For the random sample we need to compute the SE for the Latino and non-Latino parts of the population.

```{r}
se_l_a <- sqrt(p*(1 - p)/n_l_a)
se_nl_a <- sqrt(p*(1 - p)/n_nl_a)
se_l_a;se_nl_a
```

Then we compute the standard error of the comparison.

```{r}
sqrt(se_l_a^2 + se_nl_a^2)
```

For our stratified sample this is quicker, as we have already calculated the standard error for both groups.

```{r}
sqrt(se_l_b^2 + se_nl_b^2)
```

We can see that the standard error for comparison is smaller for the stratified sample.
