---
title: "Chapter 7 Exercises"
author: "JC"
format: html
editor: visual
---

```{r}
library(rosdata)
library(tidyverse)
library(rstanarm)
source("Functions/stan_glm_plot_helpers.R")
```

7.1)

It would impact our model by dampening the effect of income growth on vote share. In periods of high inflation, say the 70s, income growth increased by a large margin but would do a poor job of predicting the winner of elections. The model would in turn reduce the slope of effect from growth in personal income. Residual error would get larger as well.

7.2)

```{r}
linear_sim <- function(n = 100, a = 5,b = 7,sigma = 3){
  x <- runif(n, min = 0, max = 50)
  error <- rnorm(n, 0, sigma)
  y <- a + b*x + error
  return(data.frame(x,y))
}
```

```{r}
linear_result <- linear_sim()
```

```{r}
fit <- stan_glm(y ~ x, data = linear_result)
```

```{r}
print(fit)
```

```{r}
ggplot(data = linear_result,aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = lm) +
  annotate("text",x = 25, y = 300, label = "5.4 + 7x")
```

7.3)

```{r}
quadratic_sim <- function(n = 100, a = 5,b = 7, c = 10, sigma = 3){
  x <- runif(n, min = 0, max = 50)
  error <- rnorm(n, 0, sigma)
  y <- a + b*x + c*x^2 + error
  return(data.frame(x,y))
}
```

```{r}
quad_data <- quadratic_sim()
quad_data |> 
  ggplot(aes(x = x, y = y)) +
  geom_point()
```

a\)

```{r}
fit <- stan_glm(y ~ x, data = quad_data)
```

```{r}
grid <- tibble(x = seq(min(quad_data$x), max(quad_data$x), length.out = 1000))
fit_df <- stan_glm_fit_df(fit, grid)
```

```{r}
quad_data |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(data = fit_df, aes(y = .mean))
```

We can see that the "best linear fit", aka the one that minimizes lse does not do a good job here because the true relationship is quadratic.

7.5)

```{r}
linear_converage_sim <- function(n = 100, a = 5,b = 7,sigma = 3){
  x <- runif(n, min = 0, max = 10)
  error <- rnorm(n, 0, sigma)
  y <- a + b*x + error
  fake_data <- data.frame(x,y)
  fake_fit <- stan_glm(y ~ x, data = fake_data)
  ci <- posterior_interval(fake_fit)
  return(data.frame(ci))
}
```

```{r}
ci <- linear_converage_sim()
```

```{r}
n_seq <- c(10,30,100,300,1000,3000,10000,30000)
lapply(n_seq, function(n) linear_converage_sim(n = n))
```
