---
title: "Chapter 7 Exercises"
author: "JC"
format: html
editor: visual
---

```{r}
library(rosdata)
library(tidyverse)
library(rstanarm)
source("Functions/stan_glm_plot_helpers.R")
```

7.1)

It would impact our model by dampening the effect of income growth on vote share. In periods of high inflation, say the 70s, income growth increased by a large margin but would do a poor job of predicting the winner of elections. The model would in turn reduce the slope of effect from growth in personal income. Residual error would get larger as well.

7.2)

```{r}
linear_sim <- function(n = 100, a = 5,b = 7,sigma = 3){
  x <- runif(n, min = 0, max = 50)
  error <- rnorm(n, 0, sigma)
  y <- a + b*x + error
  return(data.frame(x,y))
}
```

```{r}
linear_result <- linear_sim()
```

```{r}
fit <- stan_glm(y ~ x, data = linear_result)
```

```{r}
print(fit)
```

```{r}
ggplot(data = linear_result,aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(method = lm) +
  annotate("text",x = 25, y = 300, label = "5.4 + 7x")
```

7.3)

```{r}
quadratic_sim <- function(n = 100, a = 5,b = 7, c = 10, sigma = 3){
  x <- runif(n, min = 0, max = 50)
  error <- rnorm(n, 0, sigma)
  y <- a + b*x + c*x^2 + error
  return(data.frame(x,y))
}
```

```{r}
quad_data <- quadratic_sim()
quad_data |> 
  ggplot(aes(x = x, y = y)) +
  geom_point()
```

a\)

```{r}
fit <- stan_glm(y ~ x, data = quad_data)
```

```{r}
grid <- tibble(x = seq(min(quad_data$x), max(quad_data$x), length.out = 1000))
fit_df <- stan_glm_fit_df(fit, grid)
```

```{r}
quad_data |> 
  ggplot(aes(x = x, y = y)) +
  geom_point() +
  geom_line(data = fit_df, aes(y = .mean))
```

We can see that the "best linear fit", aka the one that minimizes lse does not do a good job here because the true relationship is quadratic.

7.5)

```{r}
linear_converage_sim <- function(n = 100, a = 5,b = 7,sigma = 3){
  x <- runif(n, min = 0, max = 10)
  error <- rnorm(n, 0, sigma)
  y <- a + b*x + error
  fake_data <- data.frame(x,y)
  fake_fit <- stan_glm(y ~ x, data = fake_data)
  ci <- posterior_interval(fake_fit)
  return(data.frame(ci))
}
```

```{r}
ci <- linear_converage_sim()
```

```{r}
n_seq <- c(10,30,100,300,1000,3000,10000,30000)
lapply(n_seq, function(n) linear_converage_sim(n = n))
```

```{r}
election_df<- read.table(
  "Data/ROS-Examples-master/ElectionsEconomy/data/hibbs.dat", header = TRUE)
head(election_df)
```

```{r}
election_df <- election_df |> 
  mutate(high_income_growth = growth > 2)

head(election_df |> select(year,growth,high_income_growth))
```

```{r}
fit <- stan_glm(vote ~ high_income_growth, data = election_df)
```

```{r}
print(fit)
```

The effect from high income growth is larger, but so is the standard error. Using the residual error as a way of comparing the models we see that our fit got worse.

7.7)

```{r}
model_sim <- function(){
  x = runif(100,0,20)
  error = rnorm(100,0,5)
  y = 2 + 3*x + error
  fake = data.frame(x,y)
  fit <- stan_glm(y ~ x, data = fake)
  return(list(
    fake = fake,
    fit = fit
  ))
}
```

```{r}
output <- model_sim()
```

```{r}
print(output$fit)
```

```{r}
ggplot(data = output$fake, aes(x = x, y = y)) +
  geom_point()
```

```{r}
grid <- tibble(x = seq(min(output$fake$x), max(output$fake$x), length.out = 1000))
fit_df <- stan_glm_fit_df(output$fit, grid)
```

```{r}
ggplot(data = output$fake, aes(x = x, y = y)) +
  geom_point() +
  geom_line(data = fit_df, aes(y = .mean))
```

b\)

```{r}
print(output$fit)
```

The coefficients are quite close, reasonably close depends on how you view it. I think a good barometer is whether the true effect is in the 95% interval.

7.9)

```{r}
fit <- stan_glm(growth ~ vote, data = election_df)
```

```{r}
print(fit)
```

```{r}

```

```{r}
ggplot(aes(x = vote, y = growth), data = election_df) +
  geom_point() +
  geom_smooth(method = "lm")
  
```

We can see that the same relationship appears, this would be useful for predicting previous economic growth given the election result but that is about it. We wouldn't want to get confused and say that the economy is expected to grow more when votes for the incumbent party is high.
